### 2017-11-15 paper/sections/1_introduction.tex
paper: provide motivation for this work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

The fine-grained data annotation capabilities provided by key-value storage is
a natural match for many types of scientific simulation. Simulations relying on
a mesh-based decomposition of a physical region may result in millions or
billions of mesh cells. Each cell contains materials, pressures, temperatures
and other characteristics that are required to accurately simulate phenomena of
interest. In our target application, the
ParSplice~\cite{perez:jctc20150parsplice} molecular dynamics simulation, a
hierarchy of caches and a single persistent key-value store are used to store
both observed minima across a molecule's equation of motion (EOM) and the
hundreds or thousands of partial trajectories calculated each second during a
parallel job. Unfortunately, if we scale the system the IO to the storage
hierarchy will quickly saturate both the storage and bandwidth capacity of a
single node, so we need more effective data management techniques, such as
cache management or load balancing across a cluster.

>>>>

Storage systems use software-based caches extensively to improve performance
and resource utilization but the policies that guide what data to evict and
when to evict vary with the application or use case. For example, caching file
system metadata on both clients and servers reduces the number of remote
procedure calls and drastically improves the performance of create-heavy
workloads common in HPC~\cite{ren:sc2014-indexfs, patil:fast2011-giga+}. But
these techniques are specific to the domain and application's behavior and a
new use case may prove to be poor a match for the selected caching
policy~\cite{sevilla:sc15-mantle, weil:sc2004-dyn-metadata,
weil:osdi2006-ceph}. In this paper, we evaluate a variety of caching policies
using a general framework that supports policy switching and arrive at a
customized policy that works well for our target application and storage
system.  Trying general policies and iterating to a very customized solution
confirms that our prototype can adapt to different types of workloads; but more
importantly, this is the second system and domain that we have successfully used
this policy language and engine in. This breadth helps us reason about more
general data management policies by giving us a framework to compare and
contrast policies from other domains and services.

Our target application is ParSplice~\cite{perez:jctc20150parsplice}, a
molecular dynamics simulation that uses a hierarchy of caches and a single
persistent key-value store to store both observed minima across a molecule's
equation of motion (EOM) and the hundreds or thousands of partial trajectories
calculated each second during a parallel job.  The fine-grained data annotation
capabilities provided by key-value storage is a natural match for scientific
simulations like ParSplice. But these simulations rely on a mesh-based
decomposition of a physical region and result in millions or billions of mesh
cells, where each cell contains materials, pressures, temperatures and other
characteristics that are required to accurately simulate phenomena of interest.
Unfortunately, this simulations of this size saturate the capcity and bandwidth
capabilities of a single node so we need more effective data management
techniques, such as cache management or load balancing across a cluster.

### 2017-11-15 paper/sections/4_dynamic-policies.tex
paper: provide motivation for this work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
\caption{The performance and resource utilization trade-off for different cache
sizes. ``Baseline" is ParSplice unmodified and the ``Policy: Constrain Cache
Size" graph limits the size of the cache to save memory.
\label{fig:methodology-tradeoff}}
\end{figure}
%
%\footnotesize
%\centering
%\begin{minted}[xleftmargin=3em,linenos]{lua}
%function when()
%  if server[whoami]['cachesize'] > n then
%    return true
%  end
%  return false
%end
%
%function howmuch()
%  if servers[whoami]['cachesize'] > n
%    return servers[whoami]['cachesize'] - n
%  end
%  return 0
%end
%\end{minted}
%\caption{Policy that implements a basic LRU cache. The performance and memory
%utilization for different values of \(n\) are graphed in
%Figure~\ref{fig:methodology-tradeoff}.
%\label{src:lru}}
%\end{figure}

% why is the performance lower for smaller caches?
Although the keyspace grows to 150K, a 100K key cache achieves 99\% of the
performance. Decreasing the cache degrades performance and predictability.
Despite the memory savings, our results suggest that dynamic cache management
policies could save even more memory.  Figure~\ref{fig:methodology-tradeoff}
show that a 100K key cache is sufficient as a static policy but the top graph
in Figure~\ref{fig:motivation-regimes} indicates that the cache size could be
much smaller. That graph shows that the beginning of the run is characterized
by many reads to a small set of keys and the end sees much lower reads per
second to a larger keyspace. Specifically, it shows only about 100 keys as
active in the latter half of the run.

@@

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff-dynamic.png}\\
  \caption{The performance and resource utilization trade-off of using a
  dynamic cache management policy that switches to a constrained cache policy after absorbing
  the initial burstiness of the workload. The sizes of these smaller caches are
  on the \(x\) axis.  \label{fig:methodology-tradeoff-dynamic}}
\end{figure}

Figure~\ref{fig:methodology-tradeoff-dynamic} shows the results of using Mantle
to program a dynamic cache management policy into ParSplice that switches between
two policies:

@@

% results: same level of performance can be achieved 
The dynamic policies show better performance than the single \(n\) key
policies. The performance and memory utilization for a 100K key cache size is
the same as the 100K bar in Figure~\ref{fig:methodology-tradeoff-dynamic} but
the rest reduce the size of the keyspace after the read flash crowd.  We see
the worst performance when the engine switches to the 10 key limit policy,
which achieves 94\% of the performance while only using 40KB of memory.

@@

The results in Figure~\ref{fig:methodology-tradeoff-dynamic} are slightly
deceiving for three reason: (1) segments take longer to generate later in the
run, (2) the memory footprint is the value at the end of 2.5 hours, and (3)
this policy only works well for the 2.5 hour run.  For (1), the curving down of
the simulation vs. wall-clock time is shown in
Figure~\ref{fig:methodology-trajectory}; as the nanoparticle grows it takes
longer to generate segments so by the time we reach 2 hours, over 90\% of the
trajectory is already generated.  For (2), the memory footprint is around 0.4GB
until we reach 100K key switch threshold. In
Figures~\ref{fig:methodology-tradeoff}
and~\ref{fig:methodology-tradeoff-dynamic} we plot the final value. For (3),
Figure~\ref{fig:methodology-trajectory} shows that the cache fills up with 100K
keys at time 7200 seconds and its size is reduced to the size listed in the
legend.  The curves stay close to ``Unlimited" for up to an hour after the
cache is reduced but eventually flatten out as the persistent database gets
overloaded. 10K and 100K follow the ``Unlimited" curve the longest and are
sufficient policies for the 2.5 hour runs but anything longer would need a
different dynamic cache management policy.

\begin{figure}[t]
  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/methodology-trajectory.png}\\
  \caption{The rate that the trajectory is computed decays over time (which is
  expected) but this skews the performance improvements in
  Figure~\ref{fig:methodology-tradeoff-dynamic}. Our dynamic policy works for 2.5
  hour jobs but not for 4 hour jobs.  \label{fig:methodology-trajectory}}
\end{figure}

>>>>

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\

\caption{The performance and resource utilization trade-off for different cache
management policies. ``None" is ParSplice unmodified, ``Fixed Sized Cache"
limits the size of the cache by evicting the least recently used items, and
``Multi-Policy Cache" switches to a fixed sized cache after absorbing the
initial burstiness of the workload.  The size of the fixed sized cache for each
experiment is on the \(x\) axis.  \label{fig:methodology-tradeoff}}
\end{figure}
%
%\footnotesize \centering \begin{minted}[xleftmargin=3em,linenos]{lua} function
%when() if server[whoami]['cachesize'] > n then return true end return false
%end
%
%function howmuch() if servers[whoami]['cachesize'] > n return
%servers[whoami]['cachesize'] - n end return 0 end \end{minted} \caption{Policy
%that implements a basic LRU cache. The performance and memory utilization for
%different values of \(n\) are graphed in
%Figure~\ref{fig:methodology-tradeoff}.  \label{src:lru}} \end{figure}

% why is the performance lower for smaller caches?
Although the keyspace grows to 150K, a 100K key cache achieves 99\% of the
performance. Decreasing the cache degrades performance and predictability.
Despite the memory savings, our results suggest that dynamic cache management
policies could save even more memory.  The middle plot in
Figure~\ref{fig:methodology-tradeoff} show that a 100K key cache is sufficient
as a static policy but the top graph in Figure~\ref{fig:motivation-regimes}
indicates that the cache size could be much smaller. That graph shows that the
beginning of the run is characterized by many reads to a small set of keys and
the end sees much lower reads per second to a larger keyspace. Specifically, it
shows only about 100 keys as active in the latter half of the run.

@@

The plot on the right side of Figure~\ref{fig:methodology-tradeoff} shows the results of using Mantle
to program a dynamic cache management policy into ParSplice that switches between
two policies:

@@

% results: same level of performance can be achieved 
The multi-part policies show better performance than the single \(n\) key
policies. The performance and memory utilization for a 100K key cache size is
the same as the 100K bar in the ``Multi-Part Policy" graph in  Figure~\ref{fig:methodology-tradeoff} but
the rest reduce the size of the keyspace after the read flash crowd.  We see
the worst performance when the engine switches to the 10 key limit policy,
which achieves 94\% of the performance while only using 40KB of memory.

@@

The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
three reason: (1) segments take longer to generate later in the run, (2) the
memory footprint is the value at the end of 2.5 hours, and (3) this policy only
works well for the 2.5 hour run.  For (1), the curving down of the simulation
vs. wall-clock time is shown in Figure~\ref{fig:methodology-trajectory}; as the
nanoparticle grows it takes longer to generate segments so by the time we reach
2 hours, over 90\% of the trajectory is already generated.  For (2), the memory
footprint is around 0.4GB until we reach 100K key switch threshold. In
Figures~\ref{fig:methodology-tradeoff} we plot the final value. For (3),
Figure~\ref{fig:methodology-trajectory} shows that the cache fills up with 100K
keys at time 7200 seconds and its size is reduced to the size listed in the
legend.  The curves stay close to ``Unlimited" for up to an hour after the
cache is reduced but eventually flatten out as the persistent database gets
overloaded. 10K and 100K follow the ``Unlimited" curve the longest and are
sufficient policies for the 2.5 hour runs but anything longer would need a
different dynamic cache management policy.

\begin{figure}[t]
  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/methodology-trajectory.png}\\
  \caption{The rate that the trajectory is computed decays over time (which is
  expected) but this skews the performance improvements in the ``Multi-Part Policy" graph in 
  Figure~\ref{fig:methodology-tradeoff}. Our dynamic policy works for 2.5
  hour jobs but not for 4 hour jobs.  \label{fig:methodology-trajectory}}
\end{figure}
